---
title: "Handling_Missing_Values"
author: "Yu Ling Hsu"
output: html_document
Date: '2023-01-23'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data file melb_data.csv
```{r}
Melbourne_data <- read.csv("melb_data.csv") 
head(Melbourne_data)
```

Visualizing Missing Values:
```{r}
library(VIM)
library(mice)
names(Melbourne_data)

# Renaming variables to smallers size names to make them fit in the figure
names(Melbourne_data)[15] <- "BldgArea" 
names(Melbourne_data)[17] <- "CunclArea"
names(Melbourne_data)[20] <- "RgnNam"
names(Melbourne_data)[21] <- "PrpCnt"
```

Q.1 Create a new dataset (give it a new name "Melb_All_Missing_dropped") by removing all the missing values in Melbourne_data. 
Hint: Use na.omit()
```{r}
Melb_All_Missing_dropped <- na.omit(Melbourne_data)
Melb_All_Missing_dropped
```


Q.2 Calculate the percentage of datapoints dropped. 
Hint: Use nrow() to calculate the number of rows (datapoints) in a dataset.
```{r}
ndrop <- nrow(Melb_All_Missing_dropped) #put into a variable
norigin <- nrow(Melbourne_data) #put into a variable
percentage = ((norigin - ndrop) / norigin) * 100 #*100 cause want to become a precentage
percentage
#49 is more than the half, is better to use some function to keep the data
```

We see that about half (49.7%) of the data has been deleted. It can be a problem if you don't have a lot of data points.  

Now let's delete the variables that have most of the missing values. Find which variables have large number of missing values.


Q.3 Write code for visulaizing missing values using md.pattern(), aggr() and matrixplot() functions. 
Identify the variables that have most missing values. 
```{r}
md.pattern(Melbourne_data, rotate.names = TRUE) #rotate.name to see the name clearly
aggr(x = Melbourne_data)
matrixplot(x = Melbourne_data)
```
We see that the variables *BldgArea* and *YearBuilt* have large number of missing values. If these variable are not critical to our analysis, we can remove them. Let's remove them and see what percentage of data is deleted.

Q.4. Create a copy of Melbourne_data (name the copy as Melbourne_data_Var_dropped) and remove variable "BldgArea" and "YearBuilt" from it. 
```{r}
Melbourne_data_Var_dropped = Melbourne_data[,!(names(Melbourne_data) %in% c("BldgArea","YearBuilt"))]
```

Q.5. Remove all the missing values from Melbourne_data_Var_dropped and name the new dataframe as "Melb_All_Missing_dropped"
```{r}
Melb_All_Missing_dropped <- na.omit(Melbourne_data_Var_dropped)
```

Q.6. Calculate the percentage of datapoints dropped in the previous question. 
```{r}
final_persentage = ((nrow(Melbourne_data) - nrow(Melb_All_Missing_dropped)) / nrow(Melbourne_data)) * 100
final_persentage
```
We see that now only 0.45% of data is deleted.

# Instead of dropping rows with missing values, replace with mean. 
Q.7 Take the data Melbourne_data_Var_dropped (the one after dropping variables "BldgArea" and "YearBuilt"), and replace missing values in variable "Car" by its mean.
```{r}
car_mean <- mean(Melbourne_data_Var_dropped$Car, na.rm = TRUE)
car = Melbourne_data_Var_dropped$Car[is.na(Melbourne_data_Var_dropped$Car)] <- car_mean #replace NA with mean and put into variable
car
```

Q.8 Test if the data now has any missing values.
Hint: Write code for finding the total number of missing values.
```{r}
sum(is.na(Melbourne_data_Var_dropped)) #is.na to make sure if there's still missing value
```

