---
title: "Linear Regression - Covid_Twitter"
author: "Yu Ling Hsu"
date: "2/21/2022"
output: html_document
---
```{r}
##install.packages("leaps")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(leaps)
library(car)#to vif function
```


# We will use "Covid_Twitter_Sample.csv" data for this exercise. This data is a dataset about help seeking tweets during covid-19 pandemic. Each row corresponds to a particular tweet.

# Meaning of a few important variables
* reply_count: Number of replies to a tweet
* followers_count: Number of followers of the person tweeting
* friends_count: Number of people followed by the person who has tweeted
* hashtags_count: Number of hashtags in the tweet
* celebrity: Is the tweet coming from a celebrity?


Load Data
```{r}
tweets_df <- read.csv("Covid_Twitter.csv")
# Use these variables as x_variables:
x_var <- c("followers_count", "friends_count", "hashtags_count", "mentions_count", "urls_count", "media_count", "tweet_polarity", "tweet_subjectivity", "celebrity")
y_var <- 'reply_count'

length(x_var)
```


Q.1. Run a linear regression model using "reply_count" as y-variable and above mentioned x_var as x-variables.
```{r}
model <- lm(formula = paste(y_var, "~", paste(x_var, collapse = "+")), data = tweets_df)
summary(model)
```

Q.2. Write code to obtain variance inflation factor (VIF) of the above linear regression model. Does any variable have VIF higher than 4?
Hint: Use function vif()
```{r, eval = TRUE}
#no there's no varaible have VIF higher than 4
vif_values <- vif(model)
any(vif_values > 4)
```

Q.3. Run an exhaustive subset selection model on variables in "x_var". If we want to include only one x-variable in the regression, which variable should it be?
Hint: Use function regsubsets()
Use nvmax=length(x_var) to try all possible combinations of x variables. 
```{r}
model2 <- regsubsets(formula(paste(y_var, paste(x_var, collapse="+"), sep="~")), data = tweets_df, nvmax = length(x_var))
summary(model2)
```

Q.4. Obtain adjusted R-Squared values of each model in the above subset selection. 
```{r}
rsquared_values <- summary(model2)$adjr2
rsquared_values
```

Q.5. Plot adjusted R-Squared vs number of variables.
```{r}
plot(1:length(rsquared_values), rsquared_values, xlab = "Number of Variables", ylab = "Adjusted R-squared")
```

Q.6. Write code to identify the number of variables where adjusted R-Squared is maximum. 
```{r}
max_rsquared <- which.max(rsquared_values) + 1
max_rsquared
```


Q.7. According to the answer of the above question, which x-variables should not be included in the final regression (no need to write code for this)? 
Answer: 
Adjusteed R-square decreased in Number of Variables 7.This plot also confirms that using a 7-variable model is the best one.

Q.8. Plot the box-plot of the variables that should not be included in the final regression. 
```{r}
library(ggplot2)
library(reshape2)
boxplot(tweets_df$tweet_subjectivity)
boxplot(tweets_df$mentions_count)
boxplot(tweets_df$urls_count)
```


Q.9. Since "reply_count" is a count variable, run a poisson regression (between "reply_count" and variables in x_var). 
```{r}
# Fit a Poisson regression model
model <- glm(formula = paste(y_var, "~", paste(x_var, collapse = "+")), data = tweets_df, family = "poisson")
model
```